---
title: "Year-in-review report static graphs"
author: "Hope Johnson"
date: "3/3/2021"
output: html_document
---

```{r setup, include=FALSE}
##Define package list
Packages<-c("tidyverse", "skimr", "lubridate", "zoo", "plotly", "glue")
.packages = Packages
##Install CRAN packages (if not already installed)
.inst <- .packages %in% installed.packages()
if(length(.packages[!.inst]) > 0) install.packages(.packages[!.inst])
##Load packages into session 
lapply(.packages, require, character.only=TRUE)
devtools::install_github("uclalawcovid19behindbars/behindbarstools")
library(behindbarstools)

get_first_not_na <- function(x){
    if(all(is.na(x))){
        return(NA)
    }
    return(na.omit(x)[1])
}
this_dir <- file.path("~", "UCLA", "code", "year-in-review")
data_path <- file.path(this_dir, "data", "processed")
```

# Load data

Get data pre-November 11 2020, bind to with data up until February 24 2021, slightly cleaned! 

```{r include=F}
input_path <- file.path("~", "UCLA", "code", "cdc_collab", "historical-aggregate-counts")

# last updated this 2/24/21
old_scrape_df <- read_csv(file.path(input_path, "hist_df.csv")) %>%
    filter(Date <= ymd("2020-11-04"))

new_scrape_df <- read_scrape_data(TRUE) %>%
    filter((Jurisdiction == "immigration") | (Date > ymd("2020-11-04")))

comb_df <- old_scrape_df %>%
    bind_rows(new_scrape_df) %>%
    group_by(State, Name, Facility.ID, Jurisdiction) %>%
    # pass constant values to older dates where it didnt exit before
    mutate(Population.Feb20 = get_first_not_na(Population.Feb20)) %>%
    mutate(Capacity = get_first_not_na(Capacity)) %>%
    ungroup() %>%
    # only look at state and federal facilities
    filter(Jurisdiction %in% c("state", "federal", "immigration")) %>%
    # ignore statewide non facility data
    filter(!str_detect(Name, "(?i)statewide"),
           !str_detect(Name, "(?i)state-wide"))

dates <- comb_df %>%
  arrange(Date) %>%
  count(Date)
dates

ggplot(data = dates, 
       aes(x = Date, y = n)) +
  geom_bar(stat="identity") +
  labs(x = "Date",
    y = "n instances")

comb_df %>%
  count(Facility.ID)
```

Create week and 2-week averages for Residents.Confirmed. 

```{r}
## NB: only use "test_dat" for historical, facility-level vizzes when we want through-lines of data
test_dat <- comb_df %>%
  arrange(Date) %>% 
  mutate(Facility.ID = as.character(Facility.ID)) %>%
  filter(!is.na(Facility.ID)) %>% # only keep when we have a clean name
  group_by(Facility.ID) %>%
  mutate(previous_day_conf = lag(Residents.Confirmed, order_by = Date)) %>%
  mutate(conf_day_change = Residents.Confirmed - previous_day_conf) %>%
  filter(conf_day_change >= 0) %>% # filter out cumulative counts falling
  ungroup() %>%
  arrange(Facility.ID, Date) %>%
  group_by(Facility.ID) %>%
  filter(n() > 100) %>% # only keep obs when we have a through-line of historical data
  dplyr::arrange(desc(Facility.ID)) %>% 
  dplyr::group_by(Facility.ID) %>% 
  dplyr::mutate(res.confirmed_03day = zoo::rollmean(Residents.Confirmed, k = 3, fill = "extend"),
                res.confirmed_05day = zoo::rollmean(Residents.Confirmed, k = 5, fill = "extend"),
                res.confirmed_07day = zoo::rollmean(Residents.Confirmed, k = 7, fill = "extend"),
                res.confirmed_15day = zoo::rollmean(Residents.Confirmed, k = 15, fill = "extend"),
                res.confirmed_21day = zoo::rollmean(Residents.Confirmed, k = 21, fill = "extend")) %>% 
  dplyr::ungroup() %>%
  select(Facility.ID, State, Name, Jurisdiction, Date, Residents.Confirmed, 
         conf_day_change, previous_day_conf, starts_with("res.confirmed"),
         Residents.Deaths) %>%
  mutate(highlight_fac = case_when(
                Facility.ID == 100 ~ TRUE,
                Facility.ID == 652 ~ TRUE,
                Facility.ID == 83 ~ TRUE,
                Facility.ID == 112 ~ TRUE,
                TRUE ~ FALSE
          )) 
```

# Jurisidiction aggregates over time

Combine our historical data and the Marshall Project's to generate aggregate counts!

```{r}
window <- 30
collapse_vaccine <- TRUE
to_report <- c(datasets::state.name, "Federal", "ICE")
mp_data <- read_mpap_data(window = window, all_dates = TRUE) %>%
    # select(-Date) %>%
    tidyr::pivot_longer(-c(Date, State), 
                        names_to = "Measure", 
                        values_to = "MP") %>%
  filter(!is.na(Date))

state_df <- comb_df %>%
        mutate(State = ifelse(Jurisdiction == "federal", "Federal", State)) %>%
        mutate(State = ifelse(Jurisdiction == "immigration", "ICE", State)) %>%
        select(Name, State, Date, starts_with("Residents"), starts_with("Staff")) %>%
        select(-Residents.Population) %>%
        `if`(
            collapse_vaccine,
            mutate(., Staff.Vadmin = ifelse(
                is.na(.$Staff.Vadmin), .$Staff.Initiated, .$Staff.Vadmin)),
            .) %>%
        `if`(
            collapse_vaccine,
            mutate(., Residents.Vadmin = ifelse(
                is.na(.$Residents.Vadmin), .$Residents.Initiated, .$Residents.Vadmin)),
            .) %>%
        `if`(
            collapse_vaccine,
            mutate(., Staff.Vadmin = ifelse(
                is.na(.$Staff.Vadmin), .$Staff.Completed, .$Staff.Vadmin)),
            .) %>%
        `if`(
            collapse_vaccine,
            mutate(., Residents.Vadmin = ifelse(
                is.na(.$Residents.Vadmin), .$Residents.Completed, .$Residents.Vadmin)),
            .) %>%
        tidyr::pivot_longer(
            -c(Name, State, Date), names_to = "Measure", values_to = "UCLA") %>%
        filter(!is.na(UCLA)) %>%
        group_by(State, Date, Measure) %>%
        # mutate(has_statewide = c("STATEWIDE", "STATE-WIDE") %in% Name) %>%
        # # if state wide and other counts exist for a measure only use statewide
        # filter(!(has_statewide & Name != "STATEWIDE")) %>%
        summarise(UCLA = sum_na_rm(UCLA), .groups = "drop") %>%
        full_join(mp_data, by = c("State", "Date", "Measure")) %>%
        mutate(Val = case_when(
            is.na(UCLA) & is.na(MP) ~ NA_real_,
            is.na(UCLA) ~ MP,
            is.na(MP) ~ UCLA,
            UCLA >= MP ~ UCLA,
            TRUE ~ MP
        ),
        Jurisdiction = case_when(
                State == "Federal" ~ "Federal",
                State == "ICE" ~ "ICE",
                TRUE ~ "State"
              )) 

agg_df <- state_df %>%
    filter(!is.na(Val)) %>%
    group_by(Date, Measure, Jurisdiction) %>%
    summarize(
        Count = sum_na_rm(Val), 
        Reporting = sum(!is.na(Val)),
        Missing = paste0(
            to_report[!(to_report %in% State)], collapse = ", "),
        .groups = "drop") %>%
  ungroup() %>%
  mutate(week = lubridate::week(Date),
         year = lubridate::year(Date),
         week_year = glue('{week}/{year}')) 
  
rolling_max_res_confirmed <- agg_df %>%
  filter(Measure == "Residents.Confirmed") %>%
  arrange(Date) %>%
  dplyr::group_by(Jurisdiction) %>% 
  dplyr::mutate(resconfMax_07day = zoo::rollmax(Count, k = 7, fill = "extend"),
                resconfMax_10day = zoo::rollmax(Count, k = 10, fill = "extend"),
                resconfMax_14day = zoo::rollmax(Count, k = 14, fill = "extend")) %>% 
  dplyr::ungroup()

```

Create area plot showing counts aggregated by jurisdiction. 

```{r}
## RESIDENTS.CONFIRMED
#area plot
aggregate_area <- rolling_max_res_confirmed %>%
  # filter(Measure == "Residents.Confirmed") %>%
  ggplot(data = ., 
         aes(x = Date, 
             y = resconfMax_10day, 
             fill  = Jurisdiction)) +
  geom_area(alpha=1, size=.05, colour="black") +
  scale_x_date(date_breaks = "1 months", date_labels = "%b '%y") + 
  scale_y_continuous(labels = scales::comma) + 
  labs(x = "", 
       y = "Cumulative COVID-19 Cases Among Incarcerated People", 
       title = "Aggregating COVID-19 Case Counts in Prison") + 
  theme_behindbars(base_size = 18) + 
  scale_fill_bbdiscrete()
ggsave(file.path(this_dir, "plots", "aggregate_area.png"), aggregate_area, width = 14, height = 10)
  
# line plot
line <- rolling_max_res_confirmed %>%
  ggplot(data = ., 
         aes(x = Date, 
             y = resconfMax_10day, 
             group  = Jurisdiction)) +
  geom_line(aes(colour = Jurisdiction), position = "stack", alpha = 0.7, size = 0.5) + 
  scale_x_date(date_breaks = "2 months", date_labels = "%m/%y")
ggplotly(line)

## RESIDENTS.DEATHS
rolling_max_res_deaths <- agg_df %>%
  filter(Measure == "Residents.Deaths") %>%
  arrange(Date) %>%
  dplyr::group_by(Jurisdiction) %>% 
  dplyr::mutate(resdeathsMax_07day = zoo::rollmax(Count, k = 7, fill = "extend"),
                resdeathsMax_10day = zoo::rollmax(Count, k = 10, fill = "extend"),
                resdeathsMax_14day = zoo::rollmax(Count, k = 14, fill = "extend")) %>% 
  dplyr::ungroup() 

aggregate_deaths <- rolling_max_res_deaths %>%
  ggplot(data = ., 
         aes(x = Date, 
             y = resdeathsMax_10day, 
             fill  = Jurisdiction)) +
  geom_area(alpha=1, size=.2, colour="black") +
  scale_x_date(date_breaks = "1 months", date_labels = "%b '%y") + 
  scale_y_continuous(labels = scales::comma) + 
  labs(x = "", 
       y = "Cumulative COVID-19 Deaths Among Incarcerated People", 
       main = "Aggregating COVID-19 Case Counts in Prison") + 
  theme_behindbars(base_size = 18) + 
  scale_fill_bbdiscrete()
```

```{r}
# create version NOT broken down by jurisdiction
rolling_max_overall <- state_df %>%
    filter(!is.na(Val)) %>%
    group_by(Date, Measure) %>%
    summarize(
        Count = sum_na_rm(Val), 
        Reporting = sum(!is.na(Val)),
        Missing = paste0(
            to_report[!(to_report %in% State)], collapse = ", "),
        .groups = "drop") %>%
  ungroup() %>%
  mutate(week = lubridate::week(Date),
         year = lubridate::year(Date),
         week_year = glue('{week}/{year}')) %>%
  filter(Measure == "Residents.Confirmed") %>%
  arrange(Date) %>%
  dplyr::mutate(resconfMax_07day = zoo::rollmax(Count, k = 7, fill = "extend"),
                resconfMax_10day = zoo::rollmax(Count, k = 10, fill = "extend"),
                resconfMax_14day = zoo::rollmax(Count, k = 14, fill = "extend")) 

line <- rolling_max_overall %>%
  ggplot(data = ., 
         aes(x = Date, 
             y = resconfMax_14day)) +
  geom_line() + 
  scale_x_date(date_breaks = "2 months", date_labels = "%m/%y")

# write data for overall count (text element)
# rolling_max_overall %>%
#   select(Date, Reporting, starts_with("resconf")) %>% 
#   write_csv(., file.path(data_path, "agg_counts.csv"))

```


# All facilities with select highlights

```{r}
line_plot <- test_dat %>% 
  ggplot(.) + 
  ggplot2::geom_line(aes(
    x = Date, 
    y = res.confirmed_15day, # change 7-day average, 15-day average, etc here
    group = Facility.ID,
    color = highlight_fac,
    alpha = highlight_fac)) +
  # geom_line(aes(group = highlight_fac, color = "red")) +
  scale_x_date(date_breaks = "1 months", date_labels = "%b '%y") + 
  scale_y_continuous(labels = scales::comma) + 
  scale_color_manual(values=c("grey50", "red")) + 
  scale_alpha_manual(values = c(.3, 1)) + 
  labs(x = "", 
       y = "Cumulative COVID-19 Cases Among Incarcerated People", 
       title = "COVID-19 Cases In Prison") + 
  theme(legend.position = "none")
ggsave(file.path(this_dir, "plots", "facility_lines.png"), line_plot, width = 14, height = 10)


Facility_ID_lookup <- 100
test_dat %>%
  filter(Facility.ID == Facility_ID_lookup) %>%
  select(State, Name) %>%
  unique()


```
big lines:
- SUBSTANCE ABUSE TREATMENT FACILITY, CA
- AVENAL, CA (largest line)
- CALIFORNIA STATE PRISON SAN QUENTIN, CA
- Early rise: ohio, Marion 
